---
title: "Text Mining Examples"
author: "Michael Weisner"
date: "February 14, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Jane Austen Books
```{r}
library(janeaustenr)
library(stringr)
```

```{r}
tidy_books <- austen_books() %>%
  group_by(book) %>%
  mutate(linenumber = row_number(),
         chapter = cumsum(str_detect(text, regex("^chapter [\\divxlc]", ignore_case = TRUE)))) %>%
  ungroup() %>%
  unnest_tokens(word, text)

tidy_books
```

### Sentiment

There are several sentiment analysis datasets, for example here are sentiment assignments from the [National Resource Council of Canada's Emotional Lexicon](https://www.nrc-cnrc.gc.ca/eng/solutions/advisory/emotion_lexicons.html)

Let's find words that are associated with being joyful
```{r}
nrc_joy <- get_sentiments("nrc") %>% 
  filter(sentiment == "joy") # extract joyful words as determined by the nrc group

head(nrc_joy)
```

Let's look at the sentiment of words in the book "Emma"
```{r}
tidy_books %>%
  filter(book == "Emma") %>%
  inner_join(nrc_joy) %>%
  count(word, sort = TRUE)
```

Now let's do an inner join (so just of the words that are in both datsaets) of Emma with the Bing sentiment analysis
```{r}
jane_austen_sentiment <- tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(book, index = linenumber %/% 80, sentiment) %>%
  spread(sentiment, n, fill = 0) %>%
  mutate(sentiment = positive - negative)

head(jane_austen_sentiment)
```

Notice it has both negative and positive scores and a net sentiment (representing positive and negative language, for whatever that's worth).

And now let's plot it
```{r}
ggplot(jane_austen_sentiment, aes(index, sentiment, fill = book)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~book, ncol = 2, scales = "free_x")
```

Lastly, let's do a sentiment analysis of the most frequent words' contribution to positive and negative sentiment.
```{r}
tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort = TRUE) %>%
  ungroup() %>%
  group_by(sentiment) %>%
  top_n(10) %>%
  ungroup() %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(word, n, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~sentiment, scales = "free_y") +
  labs(y = "Contribution to sentiment",
       x = NULL) +
  coord_flip()  
```


### Word Clouds
R Also has good libraries for wod clouds (which are less useful for statistics, but fun)
```{r}
library(wordcloud)

tidy_books %>%
  anti_join(stop_words) %>%
  count(word) %>%
  with(wordcloud(word, n, max.words = 100))
```

We could color the words by sentiment.
A good list of colors is available [here](http://www.stat.columbia.edu/~tzheng/files/Rcolor.pdf)
```{r}
sentiment_books <- tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort = TRUE) %>%
  mutate(word = reorder(word, n))

n <- nrow(sentiment_books)
colors <- rep("grey", n)
colors[sentiment_books$sentiment == "negative"] <- "coral2"
colors[sentiment_books$sentiment == "positive"] <-  "cyan3"

sentiment_books %>%
with(wordcloud(word, n, colors = colors, max.words = 100))
```
